### 2014-08-06 Pero_Transcriptome.tex
writing pop genomics

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

@@

The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers recommendations. The error corrected adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}. Specifically, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. Reads from each tissue were assembled using Trinity version released 17 July 2014 \cite{Haas:2013jq}. We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt. The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, I downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}), and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). I used a relaxed blastN procedure (evalue set to 10$^-10$) to identify contigs in the \textit{P. eremicus} dataset that are likely biological in origin. I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity. The resulting assembly file was characterized using the software package transrate (\url{https://github.com/Blahah/transrate}). \\

@@

To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a resent work on desert-adapter \textit{Dipodomys} rodents. The coding sequence corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}) implemented in transrate. These alignments were inputted into the software codeABC version 1.6.0 \cite{Lopes:2013fd}.

@@

Transcriptome assembly for each tissue type was accomplished using the program Trinity. The raw assembly for brain, liver, testes and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively. This assembly was filtered using a blastN procedure against the \textit{Mus} cDNA and ncRNA which resulted in a final dataset containing 68331 brain-specific transcripts, 71041 liver-specific transcripts, 67340 testes-specific transcripts, and 113050 kidney-specific transcripts. Mapping the error-corrected adapter/quality trimmed reads to these datasets resulted in mapping 94.98\% (87.01\% properly paired) of brain-derived reads to the brain transcriptome, 96.07\% (88.13\% properly paired) of liver-derived reads to the liver transcriptome, 96.81\% (85.10\% properly paired) of testes-derived reads to the testes transcriptome, and 91.87\% (83.77\% properly paired) of kidney-derived reads to the kidney transcriptome. Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality.  \\

@@

\subsection*{Subsection 2}

>>>>

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing
\newcommand{\peer}{\textit{P. eremicus}}

@@

The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers recommendations. The error corrected adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}. Specifically, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. Reads from each tissue were assembled using Trinity version released 17 July 2014 \cite{Haas:2013jq}. We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt. The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM. To filter the raw sequence assembly, I downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}), and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}). I used a relaxed blastN procedure (evalue set to 10$^-10$) to identify contigs in the \textit{P. eremicus} dataset that are likely biological in origin. This procedure, when a reference dataset is available, is superior to a strategy employing expression-based filtering of the raw assembly. I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity. The resulting assembly file was characterized using the software package transrate (\url{https://github.com/Blahah/transrate}). \\

@@

To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a resent work on desert-adapter \textit{Dipodomys} rodents. The coding sequence corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapens} identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}). An unrooted gene tree was constructed using the online resource Clustal-Omega, and together the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}.

@@

Transcriptome assembly for each tissue type was accomplished using the program Trinity \cite{Haas:2013jq}. The raw assembly for brain, liver, testes and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively. This assembly was filtered using a blastN procedure against the \textit{Mus} cDNA and ncRNA which resulted in a final dataset containing 68331 brain-specific transcripts, 71041 liver-specific transcripts, 67340 testes-specific transcripts, and 113050 kidney-specific transcripts. Mapping the error-corrected adapter/quality trimmed reads to these datasets resulted in mapping 94.98\% (87.01\% properly paired) of brain-derived reads to the brain transcriptome, 96.07\% (88.13\% properly paired) of liver-derived reads to the liver transcriptome, 96.81\% (85.10\% properly paired) of testes-derived reads to the testes transcriptome, and 91.87\% (83.77\% properly paired) of kidney-derived reads to the kidney transcriptome. Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality.  \\

@@

As detailed above, the RNAseq data from 15 individuals were mapped to the reference transcriptome with the resulting BAM files being used as input to the sortware package ANGSD. The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals. The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 1), with a much smaller fraction being subject to neutral or

To test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, I implemented the branch-site test as described above, setting the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions. The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0), indicating enhanced selection in \peer\: relative to the other lineages. The branch site test for positive selection conducted on the Vdr gene was non-significant (2$\Delta$Lnl=0.68, df=1, p=1).
